---
title: "3_decontam"
output: html_document
date: "2023-06-19"
---

https://benjjneb.github.io/decontam/vignettes/decontam_intro.html

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("decontam")
```{r libraries}
library(phyloseq)
library(decontam)
library(tidyverse)
library(DT)
library(ggplot2); packageVersion("ggplot2")
```

#Identify Contaminants - Frequency
```{r Phyloseq}
ps <-readRDS("~/Documents/Lola/Datos_CIPF_2023/Rats_HE_HiperAmmon_model/outputs/ps.Rds")  
df <- as.data.frame(sample_data(ps)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(ps)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Sample_type)) + geom_point()

contamdf.freq <- isContaminant(ps, method="frequency", conc="Quant_DNA")
head(contamdf.freq)
saveRDS(contamdf.freq, "~/Documents/Lola/Datos_CIPF_2023/Rats_HE_HiperAmmon_model/outputs/contaminant_table_freq.Rds")
table(contamdf.freq$contaminant)
head(which(contamdf.freq$contaminant))
```
219  405  519  747  755 1053 1127 1465
In our phyloseq object, "Sample_type" is the sample variable that holds the negative control information. We’ll summarize that data as a logical variable, with TRUE for control samples, as that is the form required by isContaminant.
#Identify Contaminants - Prevalence
```{r prevalence}
sample_data(ps)$is.neg <- sample_data(ps)$Sample_type == "Pool_control"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg")
table(contamdf.prev$contaminant)
#Let’s try using this more aggressive classification threshold rather than the default.
contamdf.prev05 <- isContaminant(ps, method="prevalence", neg="is.neg", threshold=0.5)
table(contamdf.prev05$contaminant)

# Make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(ps, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Sample_type == "Pool_control", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Sample_type == "Sample", ps.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                      contaminant=contamdf.prev05$contaminant)
ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")
```

The dashed black line of the plots shows the model of a noncontaminant sequence feature for which frequency is expected to be independent of the input DNA concentration. The red line shows the model of a contaminant sequence feature, for which frequency is expected to be inversely proportional to input DNA concentration, as contaminating DNA will make up a larger fraction of the total DNA in samples with very little total DNA. 
```{r plots}
#Identificados por prev y manualmente
plot_frequency(ps, taxa_names(ps)[c(51,105, 851, 914, 1421, 1422)], conc="Quant_DNA") + 
    xlab("DNA Concentration measured by Qubit")
View(tax_table(ps)[c(51,105, 851, 914, 1421, 1422)])

#Identificados por freqencia
set.seed(321)
plot_frequency(ps, taxa_names(ps)[sample(which(contamdf.freq$contaminant))], conc="Quant_DNA") +
    xlab("DNA Concentration (PicoGreen fluorescent intensity)")

View(tax_table(ps)[c(219,  405,  519,  747,  755, 1053, 1127, 1465)])

```

Now that we have identified likely contaminants, let’s remove them from the phyloseq object:
```{r phyloseq 2}
ps.noncontamfreq <- prune_taxa(!contamdf.freq$contaminant, ps)
ps.noncontamfreq
saveRDS(ps.noncontamfreq, "~/Documents/Lola/Datos_CIPF_2023/Rats_HE_HiperAmmon_model/outputs/ps.noncontamfreq.Rds")  
```

```{r phyloseq 3}
ps.noncontamprev <- prune_taxa(!contamdf.prev05$contaminant, ps)
ps.noncontamprev
saveRDS(ps.noncontamprev, "~/Documents/Lola/Datos_CIPF_2023/Rats_HE_HiperAmmon_model/outputs/ps.noncontamprev.Rds")  
```

Using the information that the decontam package gave me, I took the tax tables of what had been identified as contaminants and I made a deep search on the habitats of these taxa. Finally, I selected which ones were correctly identifies as contaminants and now we will create a phyloseq object without the contaminant taxa.  
```{r phyloseq ok}
contaminants<-c("ASV914", "ASV1421")
goodTaxa <- setdiff(taxa_names(ps), contaminants)
ps.noncontam <- prune_taxa(goodTaxa, ps)
ps.noncontam <- subset_taxa(ps.noncontam, (Order!="Chloroplast") | is.na(Order))
ps.noncontam
saveRDS(ps.noncontam, "~/Documents/Lola/Datos_CIPF_2023/Rats_HE_HiperAmmon_model/outputs/ps.noncontam.Rds")  
```